Asteroseismology of GD358 with complex C/O core profiles

### 2016-10-05 msnew.tex
more changes ref report

The present analysis also allows us to place GD358 in the context of stellar evolution. 
According to the models, the precursors of DBs emerge from the born-again phase with 
envelopes containing a nearly uniform mixture of helium (He), carbon (C), and oxygen (O) 
out to the photosphere \citep{Dreizler98,Herwig99}. As the PG 1159 stars cool, the helium 
diffuses upward and gradually accumulates to form a chemically pure surface layer. Through 
the DBV instability strip, this process is still ongoing so that instead of a pure helium 
layer surrounding the carbon and oxygen core, one has a region where the carbon and helium 
are still mixed. This leads to a double layered structure, with the pure He surface layer 
overlying the remainder of the uniform He/C/O envelope, all above the degenerate C/O core. 
A key prediction of the diffusion models is that, for a given stellar mass, the pure He 
surface layer will steadily grow thicker as the DB star cools.  GD358 is the fourth DBV 
we can use to check this theory. The other three DBVs 
(\citet{Bischoff-Kim14,Sullivan08, Metcalfe03c}) have painted a picture qualitatively consistent with 
the diffusion calculations, with the hotter best fit models having thinner pure helium 
layers. With GD358, we seek to confirm this trend.

@@

GD358 was discovered in 1982 \citep{Winget82} and has been the target of the Whole 
Earth Telescope (WET) in 1990, 1994, 2000, and 2006 \citep{Provencal09, Kepler03,Winget94}. 
New observations presented here include 278 individual observing runs (1105.1 hrs) spanning 
2007-2016 (Table \ref{journal}). Each season of observations was obtained as part of 
multi-site WET campaigns \citep{wet90}.

Data reduction follows the steps described in \citet{Provencal12}. In brief, the new observations 
were obtained with CCD photometers at multiple sites, each with distinct effective bandpasses. 
We reduce the bandpass issues by using CCDs with similar detectors whenever possible and 
employing a red cutoff filter (BG40 or S8612) to normalize wavelength response and 
minimize extinction effects. We corrected each image for bias and thermal noise, and 
normalized by the flat field. Aperture photometry using the Maestro photometry pipeline 
described by \citet{Dalessio10} was performed on each image, covering a range of 
aperture sizes for the target and comparison stars. For each individual nightly run, we 
chose the combination of aperture size and comparison star(s) producing the highest 
signal/noise light curve.

We used the WQED pipeline \citep{wqed} to examine each light curve for photometric 
quality, remove outlying points, divide by suitable comparison stars, and correct for 
differential extinction. This observational technique is therefore not sensitive to 
oscillations longer than a few hours. The final product is a series of light curves 
with amplitude variations represented as fractional intensity (mmi). The unit is a linear 
representation of the fractional intensity of modulation (1 mmi $\approx$ 1 mmag). We present
our Fourier transforms (FTs) in units of modulation amplitude 
($1\ {\rm mma}=1 \times 10^{-3}\ {\rm ma} = 0.1 \% = 1\ {\rm ppt}$).

The final reduction step is to combine the individual light curves 
(an example is shown in Fig. \ref{tersk}) to produce complete light curves for GD358 for each 
observing season. As we do for all white dwarf pulators, we assume GD358 oscillates around 
a mean light level. This assumption allows us to assess overlapping light curves from multiple 
telescopes and identify and correct any vertical offsets. As discussed in detail in \citet{Provencal09}, we find 
no significant differences between the noise levels of amplitude spectra using: 
1) the combination of all light curves including overlapping 
segments from different telescopes, 2) the combination of light curves where we retain 
only higher signal to noise observations in overlapping segments and 3) combining all light 
curves using data weighted by telescope aperture.

@@

The complexities associated with GD358's pulsations (see Section \ref{freq}) led us to also 
re-reduce all available archival data \citep{Provencal09, Bradley04, Kepler03, Winget94, Winget82} to
insure continuity in methodology. The final result is a series of light curves 
for each observing season between 1982 and 2016. 
For the new observations, 2007 contains 8.1 hrs of observation, 2008 26.1 hrs, 2009 45 hrs, 
2010 201.3 hrs, 2011 401.1 hrs, 2012 150.5 hrs, 2013 87 hrs, 2014 184.6 hrs, 2015 55 hrs and 2016 90.6 hrs. 
Our coverage is not complete, and this incompleteness produces spectral leakage in the 
amplitude spectra. To quantify this, our standard procedure samples single sinusoids using exactly the 
same times as the original data for each season. The resulting amplitude spectrum, 
or ``spectral window'', is the pattern produced in the FT by a single frequency. 
The FTs for 2010, 2011, and 2014 are given in Fig.~\ref{fts}.

\begin{figure}
\plotone{gd358multidft}
\caption{Fourier Transforms of GD358 for the 2010, 2011, and 2014 observing seasons.
The frequency range of the observed series of $\ell=1$ modes discussed in Section \ref{analysis} 
is indicated by the left arrow (red).  Peaks below $\approx$ 2400 \muHz\ are combination 
frequencies discussed in Section \ref{mode} (right arrow, blue). 
(A color version of this figure is available in the online journal.)
\label{fts}
}
\end{figure}

@@

We use {\sl Period04} \citep{Lenz05} for Fourier analysis and nonlinear least squares fitting to 
identify statistically significant frequencies in each FT. We adopt the criterion that a peak must have 
an amplitude at least four times that of the average noise level in the given frequency range 
\citep{Provencal12}. This criterion places a 99.9\% probability that the peak represents a real signal,
and is not a result of random noise \citep{Scargle82, Provencal12}.  We define ``noise'' as the frequency-dependent 
average amplitude after prewhitening of the dominant frequencies. This is unquestionably a conservative estimate, 
as it is impractical to assume that the complete set of ``real'' frequencies are removed when determining
the noise level. This is inarguably true for GD358, where amplitude modulation is present, and the peaks above $\approx$ 2500 $\mu$Hz\ 
are combination frequencies (see Section \ref{mode}, \citet{Provencal09}).  Fig.~\ref{noise} displays the average noise 
as a function of frequency for the 2007-2016 observing seasons. A similar plot for the 
archival data is presented in Fig. 3 of \citet{Provencal09}. We also calculated Monte Carlo 
simulations using the routine provided in {\sl Period04}.  This routine generates a set of light curves using 
the original times, the fitted frequencies and amplitudes, and added Gaussian noise. A least squares fit is 
performed on each simulated light curve, with the distribution of fit parameters giving the uncertainties. 
Our Monte Carlo results are consistent with our average noise estimates.

@@

Having established our baseline noise levels, we proceed to frequency selection and identification.
Our standard frequency selection procedure identifies the largest amplitude resolved peak in the FT,
fits a sinusoid with that frequency to the data set, subtracts the fit from the light curve, 
recomputes the FT, examines the residuals, and repeates the process until no significant power 
remains. This technique is known as prewhitening. Prewhitening has inherent dangers, and must be 
employed with extreme vigilance, especially as we are aware of amplitude and/or frequency 
modulation in our data set. A detailed discussion of the prewhitening procedure and steps 
taken to minimize the effects of amplitude modulation is given in \citet{Provencal09}. Our final 
identifications of independent frequencies detected in each observing season are given in 
Table \ref{freq1} and Table \ref{freq2}.

@@

Fig.~\ref{schematic} presents a schematic representation of GD358's 
independent frequencies detected between 1982 and 2016.  
The features of asteroseismic importance are the localized bands between 900 and 2400
\muHz\ (1100 and 400 s). We interpret the bands to represent a series of modes of radial overtone $k$. The bands
themselves illustrate the second way in which GD358's frequencies vary. Each band consists of 
frequency detections from multiple observing seasons, but each band is significantly wider in 
frequency space than the frequency error of any single measurement, arguing that some process 
is acting on the frequencies. The simplest explanation is rotational splitting.  However, with the 
exception of the two highest frequency (shortest period) bands ($k=8$ and $k=9$ 
from \citet{Provencal09}), we find no clear evidence of stable multiplet structure in the frequency 
distributions of the lower frequency (longer period) bands (Fig.~\ref{rot}). This does not imply that the 
structures found by \cite{Winget94}, \cite{Kepler03}, \cite{Montgomery10} and others are not real. In 
particular, it is clear that the 1741 \muHz\ ($k=12$) mode can be explained by the oblique rotation model during 2006. However, 
previous work represents short snapshots of GD358. When examined over timescales of decades, these structures 
do not regularly recur, and so cannot be interpreted as simple rotational splitting. The lower frequency band widths 
are most likely the result of unknown processes that obscure any 
underlying signatures of simple rotational splitting. 
\begin{figure}
 \epsscale{1.0}
 \plotone{gd358k17_k8.eps}
% \plottwo{gd358multifk17.eps}{gd358multifk8.eps}
 \caption{The detailed schematic distribution of frequencies for the bands near 1230 ({\em{right}}), 
1300({\em{middle}}), and 2360 \muHz\ ({\em{left}}). As an example of the higher frequency bands, the 1230 and 
1300 \muHz\ bands  show no multiplet structure, unlike the 2360 \muHz\ band. Here we see clear evidence of multiplet 
 structure. Assuming the triplet represents
 $\ell=1$, this implies a rotation period of $\approx 1.5$ days. Note changes in the x scale for both axes.
 (A color version of this figure is available in the online journal.)
 \label{rot}}
\end{figure}

@@

\citet{Mont2016} recently outlined a possible explanation for the behavior seen in the low frequency modes
of these cool DAVs that should also apply to GD358. White dwarfs pulsate in nonradial g-modes. As a DBV such as 
GD358 pulsates, it experiences local surface temperature variations as large as 3000 K.  The temperature 
variation affects different modes in different ways.  An appropriate analogy is to consider each mode's 
propagation region as a box with a lid. The box is defined as the region where the mode frequency is 
less than both the buoyancy (Brunt-V{\"a}is{\"a}l{\"a}) and acoustic (Lamb) frequencies.  The box's lid 
represents the mode's outer turning point. For short period modes (such as $k=8$ and $k=9$ in GD358), the 
lid is defined by the acoustic frequency, which is relatively insensitive to surface temperature variations. 
As the star pulsates, the box will not change, so these modes should be stable.  For the longer period modes, 
the lid of the box is defined by the buoyancy frequency, which goes to zero at the base of the surface 
convection zone. This is the important point: for longer period modes, the box lid
is actually defined by the base of the convection zone, which is very sensitive to local temperature variations.  
As GD358 pulsates, its convection zone deepens and thins in response to the local temperature variations. 
In our analogy, the lid of the box moves, effectively changing the characteristics of the box. Long period modes with 
outer turning points defined by the base of the convection zone should be perturbed and this is indeed 
what we see in GD358.

@@

Finally, given the $l$=1 identification and the lack of definitive multiplet structure for the lower 
frequency modes, the best way to determine the central frequency for each band is simply to 
average the detected frequencies in the bands.  We experimented with numerous weighting techniques, 
and determined there is no significant difference in our solutions.  Our final frequencies for each band 
are given in Table~\ref{gd}. We use these periods in our asteroseismic fitting.

>>>>

The present analysis also allows us to place GD358 in the context of stellar evolution. 
According to the models, DB progenitors emerge from the born-again phase with 
envelopes composed of a nearly uniform mixture of helium (He), carbon (C), and oxygen (O) 
out to the photosphere \citep{Dreizler98,Herwig99}. As these stars cool, the mixed helium 
diffuses upward and begins to form a chemically pure helium surface layer. The resulting
expected structure is double layered, with a pure He surface layer overlying a uniformly 
mixed He/C/O envelope, all above the degenerate C/O core. A key prediction of diffusion models 
is that, for a given stellar mass, the surface He surface layer will grow thicker as the DB 
cools. The DBVs provide an important test of this prediction: we should find an increase in 
helium layer thickness for DBVs across the helium instability strip.  DBV GD358 is the fourth 
DBV we can use to check this theory. The other three DBVs 
(\citet{Bischoff-Kim14,Sullivan08, Metcalfe03c}) have painted a picture qualitatively consistent with 
the diffusion calculations, with the hotter best fit models having thinner pure helium 
layers. With GD358, we seek to confirm this trend.

@@

GD358 was discovered in 1982 \citep{Winget82} and has been the target of the Whole 
Earth Telescope (WET) in 1990, 1994, 2000, and 2006 \citep{Provencal09, Kepler03,Winget94}. 
New observations presented here include 278 individual observing runs (1195.1 hrs) spanning 
2007-2016 (Table \ref{journal}). Each season of observations was obtained as part of 
multi-site WET campaigns \citep{wet90}.

Data reduction follows the steps described in \citet{Provencal12}. In brief, the new observations 
were obtained with CCD photometers placed at multiple sites.  Each photometer has its own 
distinct effective bandpasses, which could influence the observed pulsation amplitudes. We strive 
to reduce these bandpass issues by using CCD photometers with similar detectors whenever possible and 
employing a red cutoff filter (BG40 or S8612) to normalize wavelength response and 
minimize extinction effects. We corrected each image for bias and thermal noise, and 
normalized by the flat field. We performed aperture photometry on each image 
using the Maestro photometry pipeline described by \citet{Dalessio10}. Maestro automatically covers a range of 
aperture sizes for the target and comparison stars. For each individual nightly run, we 
chose the combination of aperture size and comparison star(s) producing the highest 
signal/noise light curve.

We used the WQED pipeline \citep{wqed} to examine each light curve for photometric 
quality, remove outlying points, divide by suitable comparison stars, and correct for 
differential extinction. Since we rely on relative photometry by dividing by comparison stars,
our observational technique is not sensitive to oscillations with periods longer than a few hours. 
The final product is a series of light curves with amplitude variations represented as fractional intensity (mmi). 
The unit is a linear representation of the fractional intensity of modulation (1 mmi $\approx$ 1 mmag). We present
our Fourier transforms (FTs) in units of modulation amplitude 
($1\ {\rm mma}=1 \times 10^{-3}\ {\rm ma} = 0.1 \% = 1\ {\rm ppt}$).

Our final reduction step is to combine the individual light curves 
(an example is shown in Fig. \ref{tersk}) to create complete light curves for GD358 for each 
observing season. As we do for all white dwarf pulators, we assume GD358 oscillates around 
a mean light level. This assumption allows us to assess overlapping light curves from multiple 
telescopes and identify and correct any vertical offsets. As discussed in detail in \citet{Provencal09}, we find 
no significant differences between the noise levels of amplitude spectra using: 
1) the combination of every light curve including overlapping 
segments from different telescopes, 2) the combination of light curves where we retain 
the high signal to noise observations in overlapping segments and 3) combining all light 
curves incorporating data weighted by telescope aperture.

@@

The complexities associated with GD358's pulsations (see Section \ref{freq}) led us to also 
re-reduce all available archival data \citep{Provencal09, Bradley04, Kepler03, Winget94, Winget82} to
insure continuity in methodology. The final result is a series of light curves 
for each observing season between 1982 and 2016. 
For the new observations, 2007 contains 8.1 hrs of observation, 2008 26.1 hrs, 2009 45 hrs, 
2010 201.3 hrs, 2011 401.1 hrs, 2012 150.5 hrs, 2013 87 hrs, 2014 184.6 hrs, 2015 55 hrs and 2016 90.6 hrs. 
We are limited to ground based facilites with inherent weather issues, so our coverage is not continuous. This 
incompleteness produces spectral leakage in the amplitude spectra. To quantify this, our standard procedure 
samples single sinusoids using the exact times as the original data for each season. The resulting amplitude spectrum, 
or ``spectral window'', is the pattern produced in the FT by a single frequency. 
The FTs for 2010, 2011, and 2014 are given in Fig.~\ref{fts}.

\begin{figure}
\plotone{gd358multidft.jpg}
\caption{Fourier Transforms of GD358 for the 2010, 2011, and 2014 observing seasons.
The frequency range of the observed series of $\ell=1$ modes discussed in Section \ref{analysis} 
is indicated by the left arrow (red).  Peaks below $\approx$ 2400 \muHz\ are combination 
frequencies discussed in Section \ref{mode} (right arrow, blue). 
(A color version of this figure is available in the online journal.)
\label{fts}
}
\end{figure}

@@

We use {\sl Period04} \citep{Lenz05} for Fourier analysis and nonlinear least squares fitting to 
identify statistically significant frequencies in each observing season. Our standard procedure is to adopt the criterion 
that any peak have an amplitude at least four times above the average noise level in the given frequency range 
\citep{Provencal12}. This criterion places a 99.9\% probability that the peak represents a real signal,
and is not a result of random noise \citep{Scargle82, Provencal12}.  We define ``noise'' as the frequency-dependent 
average amplitude after removal of the dominant frequencies. This is unquestionably a conservative estimate, 
as it is impractical to assume that the complete set of ``real'' frequencies are removed when determining
the noise level. This is inarguably true for GD358, where amplitude modulation is present, and the peaks above $\approx$ 2500 $\mu$Hz\ 
are combination frequencies (see Section \ref{mode}, \citet{Provencal09}).  Fig.~\ref{noise} displays the average noise 
as a function of frequency for the 2007-2016 observing seasons. A similar plot for the 
archival data is presented in Fig. 3 of \citet{Provencal09}. We also calculated Monte Carlo 
simulations using the routine provided in {\sl Period04}.  This routine generates a data set using 
the original times, the fitted frequencies and amplitudes, and added Gaussian noise. A least squares fit is 
performed on each simulated light curve, with the distribution of fit parameters giving the uncertainties. 
Our Monte Carlo results are consistent with our average noise estimates.

@@

Having established our baseline noise levels, we proceed to frequency selection and identification.
Our standard frequency selection procedure identifies the largest amplitude resolved peak in the FT,
fits a sinusoid with that frequency to the data set, subtracts the fit from the light curve, 
recomputes the FT, examines the residuals, and repeats the process until no significant power 
remains. This technique is known as prewhitening. Prewhitening has inherent dangers, and must be 
employed with extreme vigilance, especially as we are aware of amplitude and/or frequency 
modulation in our data set. A detailed discussion of the prewhitening procedure and steps 
taken to minimize the effects of amplitude modulation is given in \citet{Provencal09}. Our final 
identifications of independent frequencies detected in each observing season are given in 
Table \ref{freq1} and Table \ref{freq2}.

@@

Fig.~\ref{schematic} presents a schematic representation of all GD358's independent frequencies 
detected between 1982 and 2016. The figure contains all peaks that meet our detection criteria, and so
represents the natural distribution of GD358's pulsation frequencies.  
The features of asteroseismic importance are the localized bands between 900 and 2400
\muHz\ (1100 and 400 s). We interpret the bands to represent a series of modes of radial overtone $k$. The bands
themselves illustrate the second way in which GD358's frequencies vary. Each band consists of 
frequency detections from multiple observing seasons, but each band is significantly wider in 
frequency space than the frequency error of any single measurement, arguing that some process 
is acting on the frequencies. The simplest explanation is rotational splitting.  However, with the 
exception of the two highest frequency (shortest period) bands ($k=8$ and $k=9$ 
from \citet{Provencal09}), we find no clear evidence of stable multiplet structure in the frequency 
distributions of the lower frequency (longer period) bands (Fig.~\ref{rot}). This does not imply that the 
structures found by \cite{Winget94}, \cite{Kepler03}, \cite{Montgomery10} and others are not real. In 
particular, it is clear that the 1741 \muHz\ ($k=12$) mode can be explained by the oblique rotation model during 2006. However, 
previous work represents short snapshots of GD358. When examined over timescales of decades, these structures 
do not regularly recur, and so cannot be interpreted as simple rotational splitting. The widths of the lower frequency bands
are most likely the result of unknown processes that obscure any 
underlying signatures of simple rotational splitting. 
\begin{figure}
 \epsscale{1.0}
 \plotone{gd358k17_k8.eps}
% \plottwo{gd358multifk17.eps}{gd358multifk8.eps}
 \caption{The detailed schematic distribution of frequencies for the bands near 1230 ({\em{right}}), 
1300({\em{middle}}), and 2360 \muHz\ ({\em{left}}). As an example of the higher frequency bands, the 1230 and 
1300 \muHz\ bands  show no multiplet structure, unlike the 2360 \muHz\ band. Here we see clear evidence of multiplet 
 structure. Assuming the triplet represents
 $\ell=1$, this implies a rotation period of $\approx 1.5$ days. Note changes in the x scale for both axes.
 (A color version of this figure is available in the online journal.)
 \label{rot}}
\end{figure}

@@

\citet{Mont2016} recently presented a possible explanation for the behavior seen in the low frequency modes
of these cool DAVs that should also apply to GD358. White dwarfs pulsate in nonradial g-modes. As a DBV such as 
GD358 pulsates, it experiences local surface temperature variations as large as 3000 K.  The temperature 
variation affects different modes in different ways.  An appropriate analogy is to consider each mode's 
propagation region as a box with a lid. The box is defined as the region where the mode frequency is 
less than both the buoyancy (Brunt-V{\"a}is{\"a}l{\"a}) and acoustic (Lamb) frequencies.  The box's lid 
represents the mode's outer turning point. For short period modes (such as $k=8$ and $k=9$ in GD358), the 
lid is defined by the acoustic frequency, which is relatively insensitive to surface temperature variations. 
As the star pulsates, the box will not change, so these modes should be stable.  For the longer period modes, 
the lid of the box is defined by the buoyancy frequency, which goes to zero at the base of the surface 
convection zone. This is the important point: for longer period modes, the box lid
is actually defined by the base of the convection zone, which is very sensitive to local temperature variations.  
As GD358 pulsates, its convection zone deepens and thins in response to the local temperature variations. 
In our analogy, the lid of the box moves, effectively changing the characteristics of the box. Long period modes with 
outer turning points defined by the base of the convection zone should be perturbed and this is indeed 
what we see in GD358.

@@

Finally, since the asteroseismic fitting deals with periods, we need to determine a period for each band 
to be used in the asteroseismic fits. Given the $l$=1 identification and the lack of definitive multiplet 
structure for the lower frequency modes, the best way to determine the central period for each band is simply 
to average the detected frequencies in each band. One might question why we do not use the central components 
of the triplets for the shorter period ($k$=9 and 8) bands. The triplets do not have exactly equal splittings, 
and the central components wander in frequency over time \citep{Provencal09}. It is probable that some process, 
such as magnetic fields, is offsetting the central components. We therefore decided to treat 
all the bands with the same protocol. We experimented with numerous weighting techniques, and determined 
there is no significant difference in our solutions.  Our measured periods for each band 
are given in Table~\ref{gd}. The errors given are the uncertainties in the mean. We use these periods in 
our asteroseismic fitting.
